---
title: "Module 1 Guided Question Set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question Set Overview

The data for this question is saved as “bp.txt”. The data set contains information on theweight (in pounds) and systolic blood pressure (in millimeters of mercury, mmHg) of 26randomly selected males in the age group 25-30.

```{r data}
summary(cars)
```

## Question 1

Create a scatterplot for this data set. Be sure to label the axes and give an appropriatetitle.  Based on the appearance of the plot, does a simple linear regression appearreasonable for the data?

```{r q1, echo=FALSE}

```

## Question 2

Perform a simple linear regression for systolic blood pressure against weight. Recordthe estimated equation for the linear regression, and state the estimated values ofthe slope and intercept. Interpret the values of the estimated slope and intercept, incontext.

## Question 3

For a 30-year-old male whose weight is 200, what is his predicted systolic blood pres-sure? What is his residual?

## Question 4

Produce the ANOVA table for this linear regression.

## Question 5

What is the value ofR2? Explain how this value is found. What does this value meanin this context?

## Question 6

What is the estimated value for the standard deviation of the error terms for thisregression model, ˆσ?

## Question 7

What are the null and alternative hypotheses for the ANOVA F test?

## Question 8

Explain how the F statistic of 35.744 is found.

## Question 9

Write an appropriate conclusion for the ANOVA F test for this simple linear regressionmodel.


