---
title: "Module 3 Guided Questions"
author: "Shannon Paylor"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question Set Overview

A glass bottle manufacturing company has recorded data on the average number of defects per 10,000 bottles due to stones (small pieces of rock embedded in the bottle wall) and the number of weeks since the last furnace overhaul.  Download the dataset “defects.txt”.

```{r data, echo=FALSE}
library(tidyverse)
defects <- read_table2("defects.txt", )
```


## Question 1

Plot defects,  the  average  number  of  defects  per  10,000  bottles,  against weeks,  the number  of  weeks  since  the  last  furnace  overhaul.   Comment  on  the  appearance  of the plot.  Do any assumptions or conditions for simple linear regression appear to be violated?  If so, which ones?

```{r plot}
defects %>% ggplot(aes(x = weeks, y = defects)) + geom_point() + xlab("Weeks Since Last Furnace Overhaul") + 
  ylab("Defects per 10,000 Bottles") + 
  labs(title = "Glass Bottle Manufacturing Defects vs. Time Since Last Furnace Overhaul")
```

While these two variables clearly have a strong positive correlation, they appear to follow more of an exponential association than a linear association, violating the assumption of a linear relationship necessary for simple linear regression.

## Question 2

Create a residual plot.  Describe the appearance of the graph of residuals versus fitted values, and comment if assumptions are not met for simple linear regression.

```{r resid_plot}
lm_defects <- lm(defects ~ weeks, data = defects)
plot(lm_defects$fitted.values, lm_defects$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")
```

This plot of residuals vs. fitted values follows a clear curved shape, providing evidence of a non-linear relationship between our regressor and predictor and violating the assumption of error terms having mean 0 for each subregion of the regressor variable.

## Question 3

Based on your answers to parts 1 and 2, do we need to transform at least one of the variables?

Yes, the plots in parts 1 and 2 show evidence of non-linearity, so we need to transform at least one of the variables to have a linear relationship.

## Question 4

One of your classmates says that since she is not sure if the variance is constant, she should use the Box-Cox method to see if the response variable should be transformed first.  Do you agree with her idea?  Briefly explain.

I agree. Because we have relatively few data points, it's hard to assess from the scatterplots whether the variance is constant or not, so it's probably a good idea to look at a Box-Cox plot to see if the response variable should be transformed first.

## Question 5

Regardless of your answer to part 4, use R to produce a plot of the profile log-likelihoods for the parameter, λ, of the Box-Cox power transformation.  What transformation, if any, would you apply to the response variable?  Briefly explain.

```{r box_cox_plot}
library(MASS)
boxcox(lm_defects)
```

Because the confidence interval for $\lambda$ does not include 1, we should use a transformation in this case. Because the optimal value of $\lambda$ is close to 0, I would choose the transformation $ln(y)$ because log transformations are more easily interpreted than other transformations.

## Question 6

Apply the transformation you specified in part 5.  Then fit another simple linear regression model and produce the residual plot to assess if the assumptions are met.

```{r transform}
defects$defects_log <- log(defects$defects)
lm_defects_transf <- lm(defects_log ~ weeks, data = defects)

plot(lm_defects_transf$fitted.values, lm_defects_transf$residuals, main="Plot of Residuals against Fitted Values with Transformed Response Variable")
abline(h=0,col="red")
```

Compared to the original residual plot the transformed model's residual plot is much better and meets the assumption of mean 0 that was not satisfied initially.

## Question 7

Create an ACF plot of the residuals.  Comment if assumptions are not met for simple linear regression.

```{r acf_plot}
acf(lm_defects_transf$residuals, main="ACF of Residuals")
```

This plots meets the assumptions for simple linear regression, as no non-zero lags have significant auto-correlation. There is no evidence that the errors are correlated (which would violate the assumption of independence).

## Question 8

Create a QQ plot of the residuals.  Comment if assumptions are not met for simple linear regression.

```{r qq_plot}
qqnorm(lm_defects_transf$residuals)
qqline(lm_defects_transf$residuals, col="red")
```

There is some evidence that the errors are non-normal in the corners of the qqplot, but given that only two points deviates far from the expected line and that linear models are robust to non-normality of error terms, I don't think it's enough evidence to abandon the simple linear regression model.
