---
title: "Module 9 Guided Questions"
author: "Shannon Paylor"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

wcgs <- read_csv("wcgs.csv")
```

## Question Set Overview

The  Western  Collaborative  Group  Study  (WCGS)  is  one  of  the  earliest  studies  regarding heart disease.  Data were collected from 3154 middle-aged males in California.  Download the file “wcsg.csv” and load it into R. We will focus on predicting the likelihood of getting a heart attack based on the following predictors:

* age.  Age in years

* sbp.  Systolic blood pressure in mm Hg

* dbp.  Diastolic blood pressure in mm Hg

* ncigs.  Number of cigarettes smoked per day, on average.

The response variable is chd69, with a ‘1’ indicating the person developed coronary heart disease, and a ‘0’ indicating the person did not develop coronary heart disease.

#### Question 1

Before fitting a model, create some graphical summaries to see if there is a difference in the distributions of the predictors among people who did and did not develop coronary heart disease.  To create a boxplot, type `boxplot(age~chd69)` to see the distribution of the variable age differs between people who did and did not develop coronary heart disease.   Create  similar  boxplots  for  the  other  variables.   Which  variables  seem  to differentiate those who developed and did not develop coronary heart disease?

```{r}
boxplot(wcgs$age ~ wcgs$chd69)
boxplot(wcgs$sbp ~ wcgs$chd69)
boxplot(wcgs$dbp ~ wcgs$chd69)
boxplot(wcgs$ncigs ~ wcgs$chd69)
```

Age, sbp, and ncigs appear to differentiate those who developed coronary heart disease and those who did not. Dbp appears fairly similar between those who did and did not develop heart disease.

#### Question 2

Use  R  to  fit  the  logistic  regression  model  using  all  the  predictors  listed  above,  and write the estimated logistic regression equation.

```{r}
result <- glm(chd69 ~ age + sbp + dbp + ncigs, data = wcgs, family = "binomial")
summary(result)
```

The estimated regression equation is log-odds = $-9.12 + 0.067age + 0.019sbp + 0.008dbp + 0.024ncigs$.

#### Question 3

Interpret the estimated coefficient for ncigs in context.

The ncigs coefficient of 0.024 means that for each one unit increase in ncigs, the odds of developing heart disease increases by about 2.4%, holding all other predictors constant.

#### Question 4

What  are  the  estimated  odds  of  developing  heart  disease  for  an  adult  male  who  is 45 years old, has a systolic blood pressure of 110 mm Hg, diastolic blood pressure of 70 mm Hg, and does not smoke?  What is this person’s corresponding probability of developing heart disease?

```{r}
predict.glm(result, newdata = tibble("age" = c(45), "sbp" = c(110), "dbp" = c(70), "ncigs" = c(0)))
#gives log odds

predict.glm(result, newdata = tibble("age" = c(45), "sbp" = c(110), "dbp" = c(70), "ncigs" = c(0)), 
            type = "response") #gives probability

exp(-3.43)

exp(-3.43)/(1 + exp(-3.43))
```

The odds of this person developing heart disease is about 0.032, and the probability of them developing heart disease is about 3.1%.

#### Question 5

Carry out the relevant hypothesis test to check if this logistic regression model with the four predictors is useful in estimating the odds of heart disease.  Clearly state the null and alternative hypotheses, test statistic, and conclusion in context.

```{r}
1 - pchisq(result$null.deviance - result$deviance, 4)
```

Need a $\Delta G^2$ test of all predictor variables. Small p-value means we reject null and conclude that at least one predictor has a non-zero coefficient.

#### Question 6

Based  on  the  Wald  test,  is  diastolic  blood  pressure  a  significant  predictor  of  heart disease, when the other three predictors are already in the model?

No, from summary we can see that p-value is large for dbp.

#### Question 7

Suppose a co-worker of yours suggests fitting a logistic regression model without the two blood pressure variables.  Carry out the relevant hypothesis test to check if this model without the blood pressure variables should be chosen over the previous model with all four predictors.

```{r}
reduced_mod <- glm(chd69 ~ age + ncigs, data = wcgs, family = "binomial")

1 - pchisq(reduced_mod$deviance - result$deviance, 2)
```

Small p-value so we reject the null hypothesis that both blood pressure coefficients are zero. Should not choose this reduced model over the full.

#### Question 8

Based on all the analysis performed, which of these four predictors would you use in your logistic regression model?

I would use age, sbp, and ncigs.
